---
title: "CK20-CK7 analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Generated on: `r Sys.time()`

Import required packages
```{r}
library(tidyverse)
library(rstatix)
library(ggpubr)
library(survival)
library(survminer)
library(labelled)
library(gtsummary)
library(ggplot2)
```

Globals
```{r}
intTumorMeasurementsFn <- "./data/ck20-ck7_measurements.csv"
clinPatAnnotFn <- "./data/clin_data_annot.csv"
regressionBySlideFn <- "./data/regression_by_slide.csv"
```

Import data
```{r}
intTumorData <- read_csv(intTumorMeasurementsFn)

intTumorData <- intTumorData %>% rename(image = "Image", class = "Class", dist_to_tumborder = "Distance in um to nearest Tumor border", area_um2 = "Area µm^2")
intTumorData <- intTumorData %>% mutate(image = str_replace(image, "(_CK20-CK7.ndpi)", ""))
```

Merge with data in KS cohort studie
```{r}
# Obtain list of images and harmonize format with KS Cohort 
imageList <- intTumorData %>% select(image) %>% summarise(pad = unique(image)) 
imageList <- imageList %>% mutate(fraction = sub('.*\\-', '', pad))
imageList <- imageList %>% mutate(pad_ck = pad)
imageList <- imageList %>% mutate(pad = sub('-[^-]*$', '', pad))
imageList <- imageList %>% mutate(pad = sub('-', '/20', pad))

# Merge image list with clindata from KS cohort
clinPatAnnotData <- read.csv(clinPatAnnotFn)
clinData <- merge(imageList, clinPatAnnotData) %>% select(pad_ck, pad, fraction, id, age, sex, neoadjuvant, replacement_percent, time_fu_os, status_fu_os, time_fu_prog, status_fu_prog)
clinData <- clinData %>% mutate(id_blocks = paste0(id, "-", fraction))
head(clinData)

# Obtain %tumor cells for the slide from KS cohort and merge
regressionBySlide <- read.csv(regressionBySlideFn)
regressionBySlide <- regressionBySlide %>% mutate(id_blocks = paste0(ids, "-", blocks)) %>% filter(id_blocks %in% clinData$id_blocks)
dim(regressionBySlide)
regressionBySlideShort <- regressionBySlide %>% select(id_blocks, percents)

# Correct %tum 0% to 0.5 % (minimal)
regressionBySlideShort <- regressionBySlideShort %>% mutate(percents = ifelse(percents == 0, 0.5, percents))

clinData <- merge(clinData, regressionBySlideShort)
clinData <- clinData %>% relocate(id_blocks, .after = id) %>% relocate(percents, .after = neoadjuvant) %>% rename(tum_percent = percents)
```

Dimensions of annotated dataset
```{r}
paste("Number of images:", length(unique(intTumorData$image)))
paste("Total number of annotations per class:")
intTumorData %>% count(class)

numInteractsByImage <- intTumorData %>% filter(class == "CK20-CK7") %>% group_by(image) %>% summarise(num_interactions = n()) %>% arrange(num_interactions)
idList <- clinData %>% distinct(id, pad_ck) %>% rename(image = "pad_ck")
numInteractsByImage <- merge(numInteractsByImage, idList)
head(numInteractsByImage)

ggplot(data=numInteractsByImage, aes(x=reorder(id, -num_interactions), y=num_interactions)) +
  geom_bar(stat="identity", fill="steelblue")+
  theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1))

write.csv(numInteractsByImage, "numInterByImage.csv", row.names = F)
```
Global density of interactions in viable metastasis area
```{r}
totalTumAreaByImage <- intTumorData %>% group_by(image) %>% summarise(tum_area_mm2 = round(sum(area_um2, na.rm = T) / 10^6, 1))
head(totalTumAreaByImage)

interAreaByImage <- merge(totalTumAreaByImage, numInteractsByImage)
head(interAreaByImage)

# Add %tumor viable to interaction datac and calculate viable tumor area
interAreaByImage <- merge(interAreaByImage, select(clinData, pad_ck, tum_percent), by.x = "image", by.y = "pad_ck") %>% relocate(tum_percent, .before = num_interactions)
interAreaByImage <- interAreaByImage %>% mutate(adj_tum_area_mm2 = (tum_area_mm2 / 100 * tum_percent))
head(interAreaByImage)

# Calculate density of interaction by VIABLE area
densityInterByImage <-interAreaByImage %>% mutate(density_interacts = num_interactions / adj_tum_area_mm2)
head(densityInterByImage)

ggplot(data=densityInterByImage, aes(x=reorder(id, -density_interacts), y=density_interacts)) +
  geom_bar(stat="identity", fill="steelblue")+
  theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1))

densityInterByImage2 <- densityInterByImage %>% select(id, num_interactions, adj_tum_area_mm2, density_interacts)
write.csv(densityInterByImage2, "densityInterByImage.csv", row.names = F)
```

Obtain progress and repl cohorts and add data on defined groups
```{r}
progression_no <- c("T9401-12", "T402-13", "T6136-13", "T1195-14", "T12401-14")
progression_100d <- c("T4886-13", "T248-13", "T1565-13", "T2956-13", "T10489-13")
repl_less5 <- c("T4954-13", "T5598-14", "T10187-14", "T11944-14", "T1104-15")
repl_5_50 <- c("T4246-13", "T10250-13", "T6205-14", "T9989-14", "T7700-12") # Obs "T12503-14" not in Manuela's images
repl_more95 <- c("T8572-13", "T4041-14", "T12544-14")

densityInterByImage <- densityInterByImage %>% mutate(pad = sub('-[^-]*$', '', image)) %>% relocate(pad, .after = image)

# From progress cohort and add subgroup data
progressCohort <- densityInterByImage %>% filter(pad %in% progression_no | pad %in% progression_100d)
progressCohort <- progressCohort %>% mutate(progress = ifelse(pad %in% progression_no, "no", "100d"))

# From replacement cohort and add subgroup data
replacementCohort <- densityInterByImage %>% filter(pad %in% repl_less5 | pad %in% repl_5_50 | pad %in% repl_more95)
replacementCohort <- replacementCohort %>% mutate(repl = case_when(
  pad %in% repl_less5 ~ "<5",
  pad %in% repl_5_50 ~ "5-50",
  pad %in% repl_more95 ~ ">95"
))
```

Analysis progression cohort
```{r}
# Merge with clinical data
progressClin <- merge(select(progressCohort, -image, -pad, -tum_percent), clinData, by = "id")

# Clinical table 
progressClinTable <- progressClin %>% select(age, sex, neoadjuvant, tum_area_mm2, num_interactions, tum_percent, adj_tum_area_mm2, density_interacts, progress)
var_label(progressClinTable) <- list(age = "Age", sex = "Gender", neoadjuvant = "Neoadjuvany chemotherapy", tum_area_mm2 = "Metastasis area (mm2)", num_interactions = "Number of CK20-CK7 interactions", tum_percent = "Viable tumor (%)", adj_tum_area_mm2 = "Area of viable tumor (mm2)", density_interacts = "CK20-CK7 interactions / mm2 viable tumor", progress = "Progression group")

progressClinTable %>% tbl_summary(type = c(age = "continuous", tum_percent = "continuous"), missing = "no", by = "progress") %>% add_p()

# Summary stats by progress group
progressCohort %>% group_by(progress) %>% get_summary_stats(density_interacts, type = "full", show = c("n", "median", "mean", "min", "max", "iqr", "sd"))
# type = c(age = "continuous", time_diagn_to_op = "continuous", time_remiss_to_eus = "continuous", time_eus_to_diagn = "continuous")

# Comparisons between no progress and 100d progress groups
# Unpaired Wilcox-test
stat.test <- progressCohort %>%
  wilcox_test(
    density_interacts ~ progress, paired = FALSE) %>%
    add_significance("p")
print(stat.test)
# Add statistical test p-values
# Create the plot
bxp <- ggboxplot(
  progressCohort, x = "progress", y = "density_interacts",
  color = "progress", palette = "jco"
  )
# Add statistical test p-values
stat.test1 <- stat.test %>% add_xy_position(x = "")
p <- bxp + stat_pvalue_manual(label = "p={round(p,3)} {p.signif}",
  stat.test1, tip.length = 0) 
print(p)
# Effect sizes
eff_size <- progressCohort %>%
  wilcox_effsize(
    density_interacts ~ progress, paired = FALSE)
print(eff_size)
```

Analysis replacement cohort
```{r}
# Merge with clinical data
replacementClin <- merge(select(replacementCohort, -image, -pad, -tum_percent), clinData, by = "id")

# Clinical table 
replacementClinTable <- replacementClin %>% select(age, sex, neoadjuvant, tum_area_mm2, num_interactions, tum_percent, adj_tum_area_mm2, density_interacts, repl)
var_label(replacementClinTable) <- list(age = "Age", sex = "Gender", neoadjuvant = "Neoadjuvany chemotherapy", tum_area_mm2 = "Metastasis area (mm2)", num_interactions = "Number of CK20-CK7 interactions", tum_percent = "Viable tumor (%)", adj_tum_area_mm2 = "Area of viable tumor (mm2)", density_interacts = "CK20-CK7 interactions / mm2 of viable tumor", repl = "Replacement group")

replacementClinTable %>% tbl_summary(type = c(age = "continuous", tum_percent = "continuous"), missing = "no", by = "repl") %>% add_p()

# Summary stats by neoajuvant
replacementCohort %>% group_by(repl) %>% get_summary_stats(density_interacts, type = "full", show = c("n", "median", "mean", "min", "max", "iqr", "sd"))

# Comparisons between no progress and 100d progress groups
# Unpaired Wilcox-test
stat.test <- replacementCohort %>%
  kruskal_test(
    density_interacts ~ repl) %>%
    add_significance("p")
print(stat.test)
# Add statistical test p-values

# Pairwise comparisons
pwc <- replacementCohort %>% 
  dunn_test(density_interacts ~ repl, p.adjust.method = "BH") 
pwc

# Create the plot
bxp <- ggboxplot(
  replacementCohort, x = "repl", y = "density_interacts",
  color = "repl", palette = "jco"
  ) + 
  labs(
    subtitle = get_test_label(stat.test, detailed = TRUE)
    )
# Add statistical test p-values
#stat.test1 <- stat.test %>% add_xy_position(x = "repl")
#p <- bxp + stat_pvalue_manual(label = "p={round(p,3)} {p.signif}",
#  stat.test1, tip.length = 0) 
print(bxp)
# Effect sizes
eff_size <- replacementCohort %>%
  kruskal_effsize(
    density_interacts ~ repl)
print(eff_size)
```

# Survival analysis OS
```{r}
#patDataOS$patient_flow <- fct_relevel(patDataOS$patient_flow, "clin surveillance", "direct surgery", "neoadjuvant", "palliative", "best suppcare")
densityInterByImage <- densityInterByImage %>% mutate(interacts_cat = ifelse(density_interacts > 20, ">20", "<20"))
densityInterByImageShort <- densityInterByImage %>% select(-pad, -id)
densityClin <- merge(clinData, densityInterByImageShort, by.x = "pad_ck", by.y = "image")

# Remove from OS the cases with major responses, where the CK20+ cells are intermixed with ductular reaction - special case
densityClin <- densityClin %>% filter(!(pad %in% c("T4246/2013", "T10250/2013", "T6205/2014", "T9989/2014", "T7700/2012")))

# Determine optimal cut-off for survival by density_interacts
# OS
surv_cutpoint(densityClin, time = "time_fu_os", event = "status_fu_os", variables = c("density_interacts"))
# DFS
surv_cutpoint(densityClin, time = "time_fu_prog", event = "status_fu_prog", variables = c("density_interacts")) 

fit<- survfit(Surv(time_fu_os, status_fu_os) ~ interacts_cat, data = densityClin)
median_surv <- surv_median(fit)
median_surv_months <- median_surv %>% mutate_at(vars(median:upper), .funs = funs(. / 365.25)) %>% mutate_if(is.numeric, round, digits=1) %>% rename ('median (years)' = median)

print(median_surv_months)

#names(fit$strata) <- gsub("patient_flow=", "", names(fit$strata))
ggsurvplot(fit, data = densityClin, pval = TRUE,
           title = "Overall survival",
           xlab = "Years",
           xscale = "d_y",
           break.time.by=365.25,
           legend.title = "Density interacts",
           risk.table = TRUE,
           risk.table.y.text = FALSE, 
           risk.table.height = 0.3
  ) # https://github.com/kassambara/survminer/issues/166
```

# Survival analysis Liver specific DFS
```{r}
#patDataOS$patient_flow <- fct_relevel(patDataOS$patient_flow, "clin surveillance", "direct surgery", "neoadjuvant", "palliative", "best suppcare")
#densityInterByImage <- densityInterByImage %>% mutate(interacts_cat = ifelse(density_interacts > 200, ">200", "<200"))
#densityInterByImageShort <- densityInterByImage %>% select(-pad)
#densityClin <- merge(clinData, densityInterByImageShort, by.x = "pad_ck", by.y = "image")

fit<- survfit(Surv(time_fu_prog, status_fu_prog) ~ interacts_cat, data = densityClin)
median_surv <- surv_median(fit)
median_surv_months <- median_surv %>% mutate_at(vars(median:upper), .funs = funs(. / 365.25)) %>% mutate_if(is.numeric, round, digits=1) %>% rename ('median (years)' = median)

print(median_surv_months)

#names(fit$strata) <- gsub("patient_flow=", "", names(fit$strata))
ggsurvplot(fit, data = densityClin, pval = TRUE,
           title = "Liver specific DFS",
           xlab = "Years",
           xscale = "d_y",
           break.time.by=365.25,
           legend.title = "Density interacts",
           risk.table = TRUE,
           risk.table.y.text = FALSE, 
           risk.table.height = 0.3
  ) # https://github.com/kassambara/survminer/issues/166
```

Analysis by distance to invasion front
```{r}
intData <- intTumorData %>% filter(class == "CK20-CK7") %>% select(-area_um2)
head(intData)

boxplot(intData$dist_to_tumborder ~ intData$image)#, 
        #log = "y")
abline(h=8500)

paste("Summary of interaction distances")
summary(intData$dist_to_tumborder) # Obs, seems to be an interaction at 6 cm from the tumor border, must be an error, needs to be removed

nrow(intData %>% filter(dist_to_tumborder > 9000))
nrow(intData %>% filter(dist_to_tumborder > 8500)) 

# 8.5 cm could be initial max dist, recheck when complete data 
intData <- intData %>% filter(dist_to_tumborder < 8500)

# Show distance distribution per image
boxplot(intData$dist_to_tumborder ~ intData$image, 
        horizontal = TRUE)
hist(intData$dist_to_tumborder, breaks = 100)

# Merge clinical data to the distance distribution dataframe
distClinLong <- intData %>% rename(pad_ck = image) 
distClinLong <- merge(distClinLong, clinData, by = "pad_ck")

# Explore distribution of interaction distance to tumor border grouped by neoadjuvant chemotherapy, % tumor regression, time progression, image identifier 
boxplot(distClinLong$dist_to_tumborder ~ distClinLong$neoadjuvant, 
        horizontal = TRUE)
boxplot(distClinLong$dist_to_tumborder ~ distClinLong$tum_percent, 
        horizontal = TRUE)
boxplot(distClinLong$dist_to_tumborder ~ distClinLong$time_fu_prog, 
        horizontal = TRUE, 
        fill = "neoadjuvant")
ggplot(distClinLong, aes(dist_to_tumborder, fill=neoadjuvant)) +
  geom_boxplot()
ggplot(distClinLong, aes(x=dist_to_tumborder)) + geom_histogram(binwidth=.5, colour="black", fill="white") +
  facet_grid(neoadjuvant ~ .)
ggplot(distClinLong, aes(x=dist_to_tumborder)) + geom_histogram(binwidth=.5, colour="black", fill="white") +
  facet_grid(tum_percent ~ .)
ggplot(distClinLong, aes(x=dist_to_tumborder)) + geom_histogram(binwidth=.5, colour="black", fill="white") +
  facet_grid(pad_ck ~ .)

# Cut interaction distances to tumor boder into bins
# set up cut-off values 
breaks <- seq(0, 8500, by = 100)
# bucketing values into bins
intData$dist_to_tumborder_cat <- cut(intData$dist_to_tumborder, breaks=breaks, include.lowest=TRUE, right=FALSE, dig.lab = 4)
# inspect bins
summary(intData$dist_to_tumborder_cat)

# Obtain the total number of interactions by image and distance intervals
numInterbyImageDist <- intData %>% group_by(image, dist_to_tumborder_cat) %>% summarise(num_interactions = n())
numInterbyImageDist <- numInterbyImageDist %>% rowwise() %>% mutate(id = clinData[ clinData$pad_ck ==  image, ]$id)
numInterbyImageDist$id <- as.factor(numInterbyImageDist$id)
head(numInterbyImageDist)

ggplot(data = numInterbyImageDist, aes(x = dist_to_tumborder_cat, y = num_interactions, group = id, colour = id)) + 
  geom_line() + theme_bw() + theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Number of CK20-CK7 interactions by distance to invasion front") +
  scale_x_discrete(breaks = levels(numInterbyImageDist$dist_to_tumborder_cat)[c(T, rep(F, 9))])

# Obtain the total number of interactions by distance intervals (all images together)
numInterbyDist <- intData %>% group_by(dist_to_tumborder_cat) %>% summarise(num_interactions = n())
head(numInterbyDist)

ggplot(data = numInterbyDist, aes(x = dist_to_tumborder_cat, y = num_interactions, group = 1)) + 
  geom_line() + theme_bw() + theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Total number of CK20-CK7 interactions by distance to invasion front") +
  scale_x_discrete(breaks = levels(numInterbyImageDist$dist_to_tumborder_cat)[c(T, rep(F, 9))])

# Obtain the average number of interactions by distance intervals (all images together)
avgInterbyDist <- numInterbyImageDist %>% group_by(dist_to_tumborder_cat) %>% summarise(avg_interactions = mean(num_interactions))
head(avgInterbyDist)

ggplot(data = avgInterbyDist, aes(x = dist_to_tumborder_cat, y = avg_interactions, group = 1)) + 
  geom_line() + theme_bw() + theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Average number of CK20-CK7 interactions by distance to invasion front") +
  scale_x_discrete(breaks = levels(numInterbyImageDist$dist_to_tumborder_cat)[c(T, rep(F, 9))])
```

Correlation analysis: interactions per image <-> growth pattern 
```{r}
library("ggpubr")

allClin <- merge(select(densityInterByImage, -image, -pad, -tum_percent), clinData, by = "id")

# % replacement growth pattern with density of interactions
ggscatter(allClin, x = "replacement_percent", y = "density_interacts", 
          #add = "reg.line", conf.int = TRUE, 
          #cor.coef = TRUE, cor.method = "pearson",
          xlab = "percentage replacement growth pattern", ylab = "number of CK7-CK20 interactions/area viable tumor")

# % replacement growth pattern with neoadjuvant chemotherapy
ggscatter(allClin, x = "replacement_percent", y = "neoadjuvant", 
          #add = "reg.line", conf.int = TRUE, 
          #cor.coef = TRUE, cor.method = "pearson",
          xlab = "percentage replacement growth pattern", ylab = "neoadjuvant chemotherapy (N/Y)")
```

Waterfall plots for density of interactions
```{r}
# waterfall plot, y axis showing number or density of interactions for each patient 
o <- order(allClin$density_interacts,decreasing=TRUE,na.last=NA)
allClin.ordered <- allClin[o,]

# colours for chemotherapy Y/N
col.Water <- ifelse(allClin.ordered$neoadjuvant == "Y", "#BC5A42", "#009296")
barplot(allClin.ordered$density_interacts, col=col.Water, border=col.Water, space=0.5,
main = "Density of CK7-CK20 interactions", ylab="interactions/area viable tumor",
cex.axis=1.0, cex.lab=1.0, legend.text=c("Neoadjuvant","Chemonaive"),
args.legend=list(fill=c("#BC5A42","#009296"), border=NA, cex=0.9))

# colours for % replacement growth pattern
col2.Water <- ifelse(allClin.ordered$replacement_percent >= 85, "#BC5A42", ifelse(allClin.ordered$replacement_percent <= 5, "#009296", "#F8B195"))
barplot(allClin.ordered$density_interacts, col=col2.Water, border=col2.Water, space=0.5,
main = "Density of CK7-CK20 interactions", ylab="interactions/area viable tumor",
cex.axis=1.0, cex.lab=1.0, legend.text=c("> 85 % replacement","5 - 85 % replacement", "< 5 % replacement"),
args.legend=list(fill=c("#BC5A42","#F8B195", "#009296"), border=NA, cex=0.9))
```

Waterfall plots for time survival/progression
```{r}
# waterfall plot, y axis showing number or density of interactions for each patient 
o.os <- order(allClin$time_fu_os,decreasing=TRUE,na.last=NA)
allClin.ordered.os <- allClin[o.os,]
o.prog <- order(allClin$time_fu_prog,decreasing=TRUE,na.last=NA)
allClin.ordered.prog <- allClin[o.prog,]

# overall survival + colours for chemotherapy Y/N
col1.Water <- ifelse(allClin.ordered.os$neoadjuvant == "Y", "#BC5A42", "#009296")
barplot(allClin.ordered.os$time_fu_os, col=col1.Water, border=col1.Water, space=0.5,
main = "Overall survival", ylab="survival [days]",
cex.axis=1.0, cex.lab=1.0, legend.text=c("Neoadjuvant","Chemonaive"),
args.legend=list(fill=c("#BC5A42","#009296"), border=NA, cex=0.9))

# progression-free + colours for chemotherapy Y/N
col2.Water <- ifelse(allClin.ordered.prog$neoadjuvant == "Y", "#BC5A42", "#009296")
barplot(allClin.ordered.prog$time_fu_prog, col=col2.Water, border=col2.Water, space=0.5,
main = "Progression-free survival", ylab="time to progression [days]",
cex.axis=1.0, cex.lab=1.0, legend.text=c("Neoadjuvant","Chemonaive"),
args.legend=list(fill=c("#BC5A42","#009296"), border=NA, cex=0.9))

# overall survival + colours for % replacement growth pattern
col3.Water <- ifelse(allClin.ordered.os$replacement_percent >= 85, "#BC5A42", ifelse(allClin.ordered.os$replacement_percent <= 5, "#009296", "#F8B195"))
barplot(allClin.ordered.os$time_fu_os, col=col3.Water, border=col3.Water, space=0.5,
main = "Overall survival", ylab="survival [days]",
cex.axis=1.0, cex.lab=1.0, legend.text=c("> 85 % replacement","5 - 85 % replacement", "< 5 % replacement"),
args.legend=list(fill=c("#BC5A42","#F8B195", "#009296"), border=NA, cex=0.9))

# progression-free survival + colours for % replacement growth pattern
col4.Water <- ifelse(allClin.ordered.prog$replacement_percent >= 85, "#BC5A42", ifelse(allClin.ordered.prog$replacement_percent <= 5, "#009296", "#F8B195"))
barplot(allClin.ordered.prog$time_fu_prog, col=col4.Water, border=col4.Water, space=0.5,
main = "Progression-free survival", ylab="time to progression [days]",
cex.axis=1.0, cex.lab=1.0, legend.text=c("> 85 % replacement","5 - 85 % replacement", "< 5 % replacement"),
args.legend=list(fill=c("#BC5A42","#F8B195", "#009296"), border=NA, cex=0.9))

# overall survival + colours for density CK7-CK20 interactions
# abline (lines) mark the samples that are removed in the plots with only >5% replacement cases

boxplot(allClin$density_interacts, # ~ allClin$neoadjuvant, 
        horizontal = TRUE)
hist(subset(allClin, replacement_percent >= 5)$density_interacts, breaks = 5)

col5.Water <- ifelse(allClin.ordered.os$density_interacts >= 200, "#CA562C", ifelse(allClin.ordered.os$density_interacts >= 150, "#EDBB8A", ifelse(allClin.ordered.os$density_interacts >= 100, "#F6EDBD", ifelse(allClin.ordered.os$density_interacts >= 50, "#B4C8A8", "#008080"))))

barplot(allClin.ordered.os$time_fu_os, 
        col=col5.Water, 
        border=col5.Water, 
        space=0.5,
        main = "Overall survival\nall Cases (n=23)", 
        cex.main=0.95, 
        ylab="survival [days]",
        cex.axis=0.8, 
        cex.lab=1.0, 
        legend.text=c("> 200","150 - 200", "100 - 150", "50 - 100", "< 50"),
        args.legend=list(title = "CK20-CK7 interactions\n / mm2 viable tumor", 
                         fill=c("#CA562C","#EDBB8A", "#F6EDBD", "#B4C8A8", "#008080"), 
                         cex=0.85, 
                         bty = "n")
        )
#abline(h=c(634,2852,2691,2728,217))

col6.Water <- ifelse(subset(allClin.ordered.os, replacement_percent >= 5)$density_interacts >= 200, "#CA562C", ifelse(subset(allClin.ordered.os, replacement_percent >= 5)$density_interacts >= 150, "#EDBB8A", ifelse(subset(allClin.ordered.os, replacement_percent >= 5)$density_interacts >= 100, "#F6EDBD", ifelse(subset(allClin.ordered.os, replacement_percent >= 5)$density_interacts >= 50, "#B4C8A8", "#008080"))))

barplot(subset(allClin.ordered.os, replacement_percent >= 5)$time_fu_os, 
        col=col6.Water, 
        border=col6.Water, 
        space=0.5,
        main = "Overall survival\n>5 % Replacement (n=18)", 
        cex.main=0.95, 
        ylab="survival [days]",
        cex.axis=0.8, 
        cex.lab=1.0, 
        legend.text=c("> 200","150 - 200", "100 - 150", "50 - 100", "< 50"),
        args.legend=list(title = "CK20-CK7 interactions\n / mm2 viable tumor", 
                         fill=c("#CA562C","#EDBB8A", "#F6EDBD", "#B4C8A8", "#008080"), 
                         cex=0.85, 
                         bty = "n")
        )
#abline(h=c(634,2852,2691,2728,217))

# progression-free survival + colours for density CK7-CK20 interactions
# abline (lines) mark the samples that are removed in the plots with only >5% replacement cases

col7.Water <- ifelse(allClin.ordered.prog$density_interacts >= 200, "#CA562C", ifelse(allClin.ordered.prog$density_interacts >= 150, "#EDBB8A", ifelse(allClin.ordered.prog$density_interacts >= 100, "#F6EDBD", ifelse(allClin.ordered.prog$density_interacts >= 50, "#B4C8A8", "#008080"))))

barplot(allClin.ordered.prog$time_fu_prog, 
        col=col7.Water, 
        border=col7.Water, 
        space=0.5,
        main = "Progression-free survival\nall Cases (n=23)", 
        cex.main=0.95, 
        ylab="time to progression [days]",
        cex.axis=0.8, 
        cex.lab=1.0, 
        legend.text=c("> 200","150 - 200", "100 - 150", "50 - 100", "< 50"),
        args.legend=list(title = "CK20-CK7 interactions\n / mm2 viable tumor", 
                         fill=c("#CA562C","#EDBB8A", "#F6EDBD", "#B4C8A8", "#008080"), 
                         cex=0.85, 
                         bty = "n")
        )
#abline(h=c(261,2852,2691,2728,217))

col8.Water <- ifelse(subset(allClin.ordered.prog, replacement_percent >= 5)$density_interacts >= 200, "#CA562C", ifelse(subset(allClin.ordered.prog, replacement_percent >= 5)$density_interacts >= 150, "#EDBB8A", ifelse(subset(allClin.ordered.prog, replacement_percent >= 5)$density_interacts >= 100, "#F6EDBD", ifelse(subset(allClin.ordered.prog, replacement_percent >= 5)$density_interacts >= 50, "#B4C8A8", "#008080"))))

bpUp <- barplot(subset(allClin.ordered.prog, replacement_percent >= 5)$time_fu_prog, 
        col=col8.Water, 
        border=col8.Water, 
        space=0.5,
        main = "Progression-free survival\n>5 % Replacement (n=18)", 
        cex.main=0.95, 
        ylab="time to progression [days]",
        cex.axis=0.8, 
        cex.lab=1.0, 
        #axes = FALSE,
        ylim = c(-1000, 3500),
        yaxp = c(-1000, 3500, 9),
        legend.text=c("> 200","150 - 200", "100 - 150", "50 - 100", "< 50"),
        args.legend=list(title = "CK20-CK7 interactions\n / mm2 viable tumor", 
                         fill=c("#CA562C","#EDBB8A", "#F6EDBD", "#B4C8A8", "#008080"), 
                         cex=0.85, 
                         bty = "n")
        )
#abline(h=c(261,2852,2691,2728,217))

bp <- barplot(height = subset(allClin.ordered.prog, replacement_percent >= 5)$replacement_percent*(0-10),
        add = TRUE,
        axes = FALSE, 
        space=c(0.6,0.4),
        col="cornflowerblue",
        border="white"
        )
text(bpUp, 500, subset(allClin.ordered.prog, replacement_percent >= 5)$id, cex=0.65)
#text(bpUp, 200, subset(allClin.ordered.prog, replacement_percent >= 5)$replacement_percent, cex=0.65)
text(bp, subset(allClin.ordered.prog, replacement_percent >= 5)$replacement_percent*(0-10)+75, subset(allClin.ordered.prog, replacement_percent >= 5)$replacement_percent, cex=0.4)
text(27.5, 500, "ID")

abline(h=c(-250,-500,-750,-1000), lty=3, lwd=0.5)


points(subset(allClin.ordered.prog, replacement_percent >= 5)$replacement_percent*30)
```

Compare number of CK7-CK20 interactions in cases that showed progression after 1000 days vs no progression after 1000 days 
```{r}
# Plot progression-free days versus % replacement 
ggscatter(allClin, x = "time_fu_prog", y = "replacement_percent", color = "neoadjuvant", size = "density_interacts", alpha = 0.5) + geom_hline(yintercept = 5, linetype = "dashed") + scale_size(range = c(0.5,12))

# Comparisons between no progress and 1000d progress groups, all samples included
# Unpaired Wilcox-test
stat.test.1000d.all <- allClin %>%
  wilcox_test(
    density_interacts ~ status_fu_prog, paired = FALSE) %>%
    add_significance("p")
print(stat.test.1000d.all)
# Add statistical test p-values
# Create the plot
bxp.all <- ggboxplot(
  allClin, x = "status_fu_prog", y = "density_interacts",
  color = "status_fu_prog", palette = "jco", add = "dotplot"
  ) + ggtitle("all Cases, n=", nrow(allClin))
# Add statistical test p-values
stat.test2.all <- stat.test.1000d.all %>% add_xy_position(x = "")
p.all <- bxp.all + stat_pvalue_manual(label = "p={round(p,3)} {p.signif}",
  stat.test2.all, tip.length = 0) 
print(p.all)
# Effect sizes
eff_size.all <- allClin %>%
  wilcox_effsize(
    density_interacts ~ status_fu_prog, paired = FALSE)
print(eff_size.all)

# Comparisons between no progress and 1000d progress groups, only samples > 5% replacement
# Unpaired Wilcox-test
stat.test.1000d.repl <- subset(allClin, replacement_percent >= 5) %>%
  wilcox_test(
    density_interacts ~ status_fu_prog, paired = FALSE) %>%
    add_significance("p")
print(stat.test.1000d.repl)
# Add statistical test p-values
# Create the plot
bxp.repl <- ggboxplot(
  subset(allClin, replacement_percent >= 5), x = "status_fu_prog", y = "density_interacts",
  color = "status_fu_prog", palette = "jco", add = "dotplot"
  ) + ggtitle("Cases with > 5% replacement, n=", nrow(subset(allClin, replacement_percent >= 5)))
# Add statistical test p-values
stat.test2.repl <- stat.test.1000d.repl %>% add_xy_position(x = "")
p.repl <- bxp.repl + stat_pvalue_manual(label = "p={round(p,3)} {p.signif}",
  stat.test2.repl, tip.length = 0) 
print(p.repl)
# Effect sizes
eff_size.repl <- subset(allClin, replacement_percent >= 5) %>%
  wilcox_effsize(
    density_interacts ~ status_fu_prog, paired = FALSE)
print(eff_size.repl)
```